---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# infinitefactor

<!-- badges: start -->
<!-- badges: end -->

The goal of `infinitefactor` is to provide samplers and post-processing functions for semi-parametric Bayesian infinite factor models. Implements different priors for the factor loadings such as MGPS, Row-wise and Column-wise Dirichlet-Laplace, CUSP, SIS and CMR (TODO: include references). Additionally, contains post-processing functions to return matrices that display rotational ambiguity to identifiability through successive application of orthogonalization procedures and resolution of column label and sign switching (TODO: rewrite this and include citation).

Although `infinitefactor` can be used as a standalone tool, it's most important use is in the `blfr` package which implements Bayesian Latent Factor Regression models.

## Installation

You can install the development version of infinitefactor from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("edovt/infinitefactor")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(infinitefactor)
## basic example code
```
