---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# infinitefactor

<!-- badges: start -->

<!-- badges: end -->

The goal of `infinitefactor` is to provide samplers and post-processing functions for semi-parametric Bayesian infinite factor models and Bayesian Latent Factor Regression with Interactions from F. Ferrari & D. Dunson (2021) [(doi:10.1080/01621459.2020.1745813)](https://doi.org/10.1080/01621459.2020.1745813). Different priors for the factor loadings are implemented, such as MGPS from A. Bhattacharya & D. Dunson (2011) [(doi:10.1093/biomet/asr013)](https://doi.org/10.1093/biomet/asr013), Row-wise Dirichlet-Laplace from A. Bhattacharya et al. (2015) [(doi:10.1080/01621459.2014.960967)](https://doi.org/10.1080/01621459.2014.960967), CUSP from S. Legramanti et al. (2020) [(doi:10.1093/biomet/asaa008)](https://doi.org/10.1093/biomet/asaa008), SIS from L. Schiavon (2022) [(doi:10.1093/biomet/asab056)](https://doi.org/10.1093/biomet/asab056) and CMR from E. Bersson et al. (2025) [(doi:10.48550/ARXIV.2504.08220)](https://doi.org/10.48550/ARXIV.2504.08220). Post-processing functions include the MatchAlign algorithm from E. Poworoznek et al. (2025) [(doi:10.1214/25-BA1544)](https://doi.org/10.1214/25-BA1544), which returns identifiable factor loadings through succesive application of orthogonalization procedures and resolution of column label and sign switching.

## Installation

You can install the development version of infinitefactor from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("edovt/infinitefactor")
```

## Example - bfa

This is a basic example of Bayesian Factor Analysis:

```{r example_bfa}
library(infinitefactor)

set.seed(219)

# 0. Data simulation as in Bhattacharya & Dunson (2011)
sim <- sim_data_bfa(n = 200, p = 20, k = 3)

# 1. Adaptive number of factors, starting at floor(3*log(p)) = 8
fit_mgps <- bfa_mgps(sim$X, verbose = FALSE)
k_fit <- summary_k(fit_mgps, real = sim$k_true) # returns median too
plots_Sigma_X <- plot_Sigma_X(fit_mgps, real = sim$Sigma_X_true)
plots_Sigma_X$scatter

# 2. Fixed number of factors -> MatchAlign
# MatchAlign seems to indicate that it is enough to use k = 3, the true k
fit_mgps_fixed <- bfa_mgps(sim$X, adapt = FALSE, k_init = k_fit, verbose = FALSE)
aligned <- match_align(fit_mgps_fixed)
plot_match_align(aligned, "Lambda", "mean")
```

## Example - blfr

```{r example_blfr}
set.seed(219)

# 1. Example without interactions
sim <- sim_data_blfr(n = 200, n_test = 100, p = 20, k = 3)
fit_mgps <- blfr_mgps(sim$y, sim$X, verbose = FALSE)

# 1.1 Induced effect summaries
Sigma_X_plots <- plot_Sigma_X(fit_mgps, real = sim$Sigma_X_true)
Sigma_X_plots$scatter
main_effects_plots <- plot_effects_X(
    fit_mgps,
    real_beta = sim$beta_X
)
main_effects_plots$main

# 1.2 Match Align - seems to indicate that it is enough to use k = 3, the true k
aligned <- match_align(fit_mgps)
plot_match_align(aligned, "Lambda", "mean")

# 1.3 Estimates on test set
preds <- predict_blfr(fit_mgps, sim$X_test)
```